A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-06-02 main.tex
コメントを質問の形に書き換えた。

\newcommand{\I}{\textrm{Int}}
\newcommand{\M}{\textrm{Mat}}

@@

In \LTP, quote and unquote are written $\TB_\alpha M$ and $\TBL_\alpha M$, respectively.
The type of $\TB_\alpha M$ is $\TW_\alpha \tau$ when $M$ has $\tau$ type.
The type of $\TBL_\alpha M$ is $\tau$ when $M$ has $\TW_\alpha \tau$.
Please notice that $\TBL_\alpha M$ is well-typed, $M$ has a code type from the typing rule of \LTP.
In addition to normal $\beta$-reduction, there is an reduction rule for unquoted quoted code.
\begin{align*}
	\TBL_\alpha (\TB_\alpha M) \longrightarrow M
\end{align*}
It means when a quoted code fragment is unquoted, it returns back to the original code fragment.
\red{This reduction is called $\longrightarrow_\Lambda$ in \LMD.}

@@

The subscription $\alpha$ is a transition variable which is called a stage variable in \LMD.
It is used to show the thickness of quoting.
For example, $\TB_\alpha (\lambda x:\I.x+10)$ is a fragment of code which becomes $(\lambda x:\I.x+10)$ after run it once and
\red{$\TB_\alpha \TB_\beta (\lambda x:\I.x+10)$ becomes $(\lambda x:\I.x+10)$ after run it twice.}

@@

\blue{Should we use $\vdash^A$? or $\vdash @ A$?}

@@

For example, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ (\beta\gamma)$ reduces to $\TB_{\beta\gamma} (\lambda x:\I.x)$.
\red{Although we can apply any transition to a transition abstraction in \LTP,
we restrict the transitions only to $\epsilon$ transition in \LMD to simplify the system.}

@@

% run
The main purpose of transition abstractions and transition applications are to represent {\bf{run}} operator.
In multistage calculus, {\bf{run}} is a very important operator.
It changes quoted code to the original code.
{\bf{run}} is realized with application of the empty transition $\epsilon$.
For example, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ \epsilon$ becomes $\lambda x:\I.x$
because $\TB_\epsilon$ is disappeared.

% CSP
CSP, cross-stage persistence, is an important feature of multistage calculus.
\red{It enable us to embed value at an outer transition into an inner transition}.
$\%$ is dedicated for CSP in \LTP.
For example, $\lambda a:\I.\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x+\%_\alpha a))$

% omitting residualization

@@

\subsection{Dependent Types}

% \LLF
\LLF is a simple \red{system} of dependent types introduced in \cite{attapl}.
It is based on Edinburgh LF\cite{harper1993framework}.
Therefore, all constants and base types are declared in the signature.
The \LLF type theory generalizes simply typed lambda calculus
by replacing the function type $\tau\to\sigma$ with the dependent product type $\Pi x:\tau.\sigma$.

@@

Next, we develop \LMD by extending \LTP with \LLF-like dependent types.

>>>>

\newcommand{\I}{\textrm{Int}}
\newcommand{\B}{\textrm{Bool}}
\newcommand{\M}{\textrm{Mat}}

@@

In \LTP, quote and unquote are written $\TB_\alpha M$ and $\TBL_\alpha M$, respectively.
The type of $\TB_\alpha M$ is $\TW_\alpha \tau$ when $M$ has $\tau$ type.
The type of $\TBL_\alpha M$ is $\tau$ when $M$ has $\TW_\alpha \tau$.
Please notice that $\TBL_\alpha M$ is well-typed, $M$ has a code type from the typing rule of \LTP.
In addition to normal $\beta$-reduction, there is an reduction rule for unquoted quoted code.
\begin{align*}
	\TBL_\alpha (\TB_\alpha M) \longrightarrow M
\end{align*}
It means when a quoted code fragment is unquoted, it returns back to the original code fragment.
This reduction is called $\longrightarrow_\Lambda$ in \LMD.
\red{この文はここに来るべきなのか? LTPの説明ではないが}

@@

The subscription $\alpha$ is a transition variable which is called a stage variable in \LMD.
It is used to show the thickness of quoting.
For example, $\TB_\alpha (\lambda x:\I.x+10)$ is a fragment of code which becomes $(\lambda x:\I.x+10)$ after run it once and
$\TB_\alpha \TB_\beta (\lambda x:\I.x+10)$ becomes $(\lambda x:\I.x+10)$ after run it twice.
\red{run twiceは表現が微妙}

@@

\red{LTPに合わせて$\vdash^A$を使うべきか? それともLMDに合わせて$\vdash @ A$を使うべきか?}

@@

For example, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ (\beta\gamma)$ reduces to $\TB_{\beta\gamma} (\lambda x:\I.x)$.
Although we can apply any transition to a transition abstraction in \LTP,
we restrict the transitions only to $\epsilon$ transition in \LMD to simplify the system.
\red{この文はここに来るべきなのか? LTPの説明ではないが}

@@

% Run
The main purpose of transition abstractions and transition applications are to represent {\bf{run}} operator.
In multistage calculus, {\bf{run}} is a very important operator.
It changes quoted code to the original code.
{\bf{run}} is realized with application of the empty transition $\epsilon$.
For example, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ \epsilon$ becomes $\lambda x:\I.x$
because $\TB_\epsilon$ is disappeared.

% CSP
CSP, cross-stage persistence, is an important feature of multistage calculus.
It enable us to embed value at an outer transition into an inner transition.
\red{inner / outer は適切か?}
$\%$ is dedicated for CSP in \LTP.
For example, $\lambda a:\I.\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x+\%_\alpha a))$

% Omitting Residualization

@@

\subsection{\LLF}

% \LLF
\LLF is a simple system of dependent types introduced in \cite{attapl}.
\red{system / type system / calculus ?}
It is based on Edinburgh LF\cite{harper1993framework}.
Therefore, all constants and base types are declared in the signature.
The \LLF type theory generalizes simply typed lambda calculus
by replacing the function type $\tau\to\sigma$ with the dependent product type $\Pi x:\tau.\sigma$.

@@

Next, we develop \LMD by extending \LTP with \LLF-like dependent types.
From here, we use the word of stage instead of transition 
because we develop a multistage calculus not a multitransition calculus.
\red{stageのほうが言葉としてふさわしいと言いたい}
There are three technical points in the extension from \LTP to \LMD.

% Constants and Base Types

First, the way of handling of constants and base types is diffrence between \LMD and \LTP.
We adopt a signature $\Sigma$ to handle constants and base types.
This is because a signature simplify kinding rules relating to type variables.
A signature $\Sigma$ is composed of pairs of a base type and its kind or a constant and its type.
For example, if you want to use boolean, $\Sigma = \B::*, \text{true}:\B, \text{false}:\B$
\red{具体的な導出例を出したほうがよいか? また、なぜsimpleになるのを書くべきか?}

% Kidinding and Well-formed Kinding Rules

Second, we need kinding rules and well-formed kinding rules in order to extend \LMD with dependent types.
It was lucky that almost all rules are determined easily.
This is because multistage calculus and dependent types are almost orthogonal.
\red{orthogonal は抽象的すぎるか?}
Therefore, we can get kinding rules and well-formed kinding rules of \LMD just by 
attaching stage anotations to ones of \LLF.
For example, \KAbs-LF is a kinding rule for a dependent type in \LLF and \KAbs\ is a corresponding one.
\begin{center}
	\infrule{\G\vdash \tau :: * \andalso \G,x:\tau@A\vdash \sigma::J}{\G\vdash(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J)}{\KAbs-LF} \\[2mm]
	\infrule{\G\V \tau :: *@A \andalso \G,x:\tau@A\V \sigma::J@A}{\G\V(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J)@A}{\KAbs} \\[2mm]
\end{center}

% Equivalence Rules

Third, we also need type equivalence rules in \LMD as a dependent type system.

% Q-% rule
